Mixture-of-interests models for representing users with diverse interests

### 2017-11-08 paper/paper.tex
Add init scaling to factorization mixtures.

\section{Model}
Let $U_i$ be a $m \times k$ matrix representing the $m$ tastes of user $i$, and $A_i$ be a $m \times k$ matrix representing the affinities of each taste for representing particular items. The recommendation score for item $j$, represented by a $k times 1$-dimensional embedding vector $e_j$, is then given by
\begin{equation}
  r_{ij} = \sigma\left(A_ie_j\right) \cdot U_ie_j,
\end{equation}
where $\sigma$ is the elementwise sigmoid function, $\sigma\left(A_ie_j\right)$ gives the mixture probabilities, and $U_ie_j$ the recommendation scores output by each mixture component. We assume identity variance matrices for all mixture components.

@@

How $U_i$ and $A_i$ are obtained differs between the three evaluated models. In the Mixture-LSTM model, $U_{it}$ and $A_{it}$ (now indexed by $t$, their position in the sequence of interactions) are linear functions of the hidden state of an LSTM layer trained on the user's previous interactions, $z_{it}$:
\begin{equation}
  z_{it} = \mathrm{LSTM}\left(e_{i1}, e_{i2}, \ldots, e_{it}\right).
\end{equation}
Given $z_{it}$, the $m$-th row of $U_{it}$ and $A_{it}$ is given by
\begin{equation}
\begin{aligned}
  u^m_{it} &= W^U_mz_{it} + B^U_m\\
  a^m_{it} &= W^A_mz_{it} + B^A_m,\\
\end{aligned}
\end{equation}
where $W^U_m$, $W^A_M$, $B^U_m$, and $B^A_m$ are the learned projection matrices. These are common across all users, representing a modest increase in the total number of model parameters. Note that the LSTM network only needs to be run once to obtain the full user representation.

@@

>>>>

\section{Model}
We propose three variants of the mixture-of-tasets model, covering both recurrent models and traditional factorization models. The \textbf{Mixture-LSTM} model uses a long-short term memory network (LSTM, \cite{hochreiter1997long}) followed by linear projection layers to obtain the mixture-of-tastes representation. By using a recurrent architecture, we capture both information from both the identity of the items the user interacted with as well as the order in which those interactions occurred. The \textbf{Projection Mixture} model . It is, however, sequence-oblivious, and . The \textbf{Embedding Mixture} model differs by directly embedding all the mixture parameters for each user.

Formally, let $U_i$ be a $m \times k$ matrix representing the $m$ tastes of user $i$, and $A_i$ be a $m \times k$ matrix representing the affinities of each taste for representing particular items. The recommendation score for item $j$, represented by a $k times 1$-dimensional embedding vector $e_j$, is then given by
\begin{equation}
  r_{ij} = \sigma\left(A_ie_j\right) \cdot U_ie_j,
\end{equation}
where $\sigma$ is the elementwise sigmoid function, $\sigma\left(A_ie_j\right)$ gives the mixture probabilities, and $U_ie_j$ the recommendation scores output by each mixture component. We assume identity variance matrices for all mixture components.

@@

How $U_i$ and $A_i$ are obtained differs between the three evaluated models. In the Mixture-LSTM model, $U_{it}$ and $A_{it}$ (now indexed by $t$, their position in the sequence of interactions) are linear functions of the hidden state of an LSTM layer trained on the user's previous interactions, $z_{it}$:
\begin{equation}
  \label{eq:lstm}
  z_{it} = \mathrm{LSTM}\left(e_{i1}, e_{i2}, \ldots, e_{it}\right).
\end{equation}
Given $z_{it}$, the $m$-th row of $U_{it}$ and $A_{it}$ is given by
\begin{equation}
\begin{aligned}
  u^m_{it} &= W^U_mz_{it} + B^U_m\\
  a^m_{it} &= W^A_mz_{it} + B^A_m,\\
\end{aligned}
\end{equation}
where $W^U_m$, $W^A_M$, $B^U_m$, and $B^A_m$ are the learned projection matrices. These are common across all users, representing a modest increase in the total number of model parameters. Note that the LSTM network only needs to be run once to obtain the full user representation.

@@

\subsection{Baselines}
Our baselines are exact equivalents of the mixture models, differing only in the fact that they represent the user with a single $k$-dimensional vector. For sequence-based models, we use an LSTM architecture and represent the user directly with the last hidden state of the network ($z_{it}$ from equation \ref{eq:lstm}). For factorization models, we use a standard latent embedding vector, corresponding directly to $z_i$ from the projection mixture model. This makes the baselines model particularly suitable for evaluating mixture-of-tastes representations: adding multiple tastes is the only source of any performance differences between the models.
